Layered Formal Verification of a TCP Stack

### 2021-06-11 SecDev21/paper.tex
Final English review by Ben

\author{%
\IEEEauthorblockN{Guillaume Cluzel}
\IEEEauthorblockA{\textit{AdaCore \& ENS de Lyon}}
\and
\IEEEauthorblockN{Kyriakos Georgious}
\IEEEauthorblockA{\textit{AdaCore \& University of Bristol}}
\and
\IEEEauthorblockN{Yannick Moy}
\IEEEauthorblockA{\textit{AdaCore}}
\and
\IEEEauthorblockN{Clément Zeller}
\IEEEauthorblockA{\textit{Oryx Embedded}}
}

@@

By 2030 it is estimated that one trillion Internet of Things (IoT) devices will be deployed and be part of every aspect of our daily life~\cite{IoT_route:ARM_white_paper}. At the same time, it is made evident that security for such systems is a major challenge~\cite{Sec_IoT_Challenges:paper,Cyber_Systems_security:paper}. Any vulnerability in software libraries that are widely used by embedded systems can potentially open the door for compromising the security of billions of deployed IoT devices. Researchers have shown that such vulnerabilities exist: in 2019, critical security vulnerabilities were discovered in a TCP implementation that has been used by billions of IoT devices for more than 13 years~\cite{zero_days_vuln:ARMIS_white_paper}, some of them allowing remote code execution.

Networking communication models define the processes of transferring information from one network component to another. Currently, the two most popular networking communication models are the TCP/IP (Transmission Control Protocol/Internet Protocol) \cite{TCP_IP:ietf_tutorial} and the OSI (Open System Interconnection) \cite{ISO:35.100} models. Layering is an essential design requirement for both models to achieve modularity, flexibility, and abstraction. Applications that only need to use the lower levels' functionality can drop all the unnecessary upper levels. \Cref{Fig:TcpStack} shows the various layers of the OSI model. Each layer supports a wide range of protocols that are compatible with the layer's specifications. The lowest layers, bottom layers in \Cref{Fig:TcpStack}, are more hardware-oriented, while the top layers are more software-specific. While the \emph{application} level has a wide range of protocols that can be used depending on the application specification, the \emph{transport} layer is heavily dependent on the two dominant protocols: Transmission Control Protocol (TCP) and User Datagram Protocol (UDP). Today, almost every networking application, including IoT devices, is using either the TCP or the UDP protocols to traffic data between the Network and the Application layers. %Thus, the reliability, assurance, and security of these protocols are of paramount importance for critical applications.

@@

In this paper, we take a practical approach, where the security of an existing TCP/IP stack implementation written in the C programming language is incrementally enhanced by replacing parts of its code with formally verified code in SPARK~\cite{mccormick_chapin_2015}, a subset of the Ada programming language supported by formal verification tools. We chose CycloneTCP, a professional-grade open source embedded TCP/IP library developed by the Oryx Embedded company~\cite{CycloneTCP}. \Cref{Fig:TcpStack} shows the protocols supported by the library. This work focuses on the TCP and Socket components of the stack. The library's TCP implementation is meant to conform to the RFC 793 protocol specifications~\cite{rfc793}. The quality of the CycloneTCP is acknowledged by the AMNESIA:33 report~\cite{AMNESIA33}, which classifies it as one of the most resilient TCP/IP stack.

In total, for this work, 50 subprograms have been translated from C to SPARK. This accounts for 2266 lines of code, of which 1165 are SPARK logic code (e.g.\ contracts) used to represent and prove TCP specification. The overall proving time with GNATprove from the Community 2021 release is 10 minutes on a 12-processor Intel(R) Core(TM) i7-9750H CPU with 2.60GHz speed and 16GB of RAM. The tools and code are freely available online.

The contributions of this article are the following. First, we show that formal verification can be applied selectively to a layer of an existing TCP/IP library, by translating the TCP layer to SPARK and modelling the lower layers on which it depends. The benefits are three-fold: we found two bugs in the TCP layer, we proved the correction of that layer, and we defined formal contracts for the upper layers to call the TCP layer. Second, we showed how the models of the lower layers can be validated by symbolic execution with KLEE on the C code of those layers.

@@

TCP is based on a multitasking model. Different tasks can interact and update the socket data structure, used to retain the status of a TCP connection, to handle the various events that can occur within a TCP session. RFC 793, under Section ``3.9. Event Processing'' page 52, describes a possible implementation of how to handle these events based on three tasks: one for the \emph{user calls}, one for the \emph{arriving segments} and one for the \emph{timers}. The role of each of the three tasks can be summarized as follows. User calls refer to functions, namely \texttt{open}, \texttt{close}, \texttt{abort}, \texttt{send} and \texttt{receive}, that can be called by the user to control the connection, send, or receive data. These functions can trigger transitions between the connection's states since they are intended to control the connection. In the arriving segments task, the received segments are being processed,
and the corresponding messages are sent back. Transitions between states can be triggered on the reception of a segment. Finally, timers control the timeouts, for example, the retransmission timeout to retransmit a message or the time-wait timeout to close the connection after a specified amount of time elapsed. Thus, transitions corresponding to the timeout events can also be triggered by this task. A complete description of TCP specification can be found in the RFC 793 document~\cite{rfc793}.

@@

Ada is a general-purpose, procedural programming language that puts great emphasis on the safety and correctness of the program. Some of the safety characteristics of the language are its strong typing system and its extensive compile-time and runtime checks. Common C programming vulnerabilities like using unsafe pointers or improper null termination of strings cannot exist in an Ada program, while issues like buffer overflow or integer overflow can be dynamically captured by Ada’s runtime checks and dealt with via exception handlers. Furthermore, Ada introduced contracts, such as preconditions and postconditions, as part of the language’s standard syntax. Such contracts are vital for explicitly expressing software and verification requirements in the source code to allow for static analyzers or runtime checks to verify that the stated requirements are met.

By subsetting Ada, the SPARK programming language \cite{mccormick_chapin_2015} emerged to provide the largest possible subset of Ada, which is suitable for functional specification and static verification. For instance, in SPARK, expressions should be free from side effects, and aliasing is forbidden (no two variables should share the exact memory location or overlap in memory), including when using pointers thanks to the use of an ownership policy~\cite{dross2020recursive}.

\emph{GNATprove} \cite{GNATProve:users_manual} is the tool that provides formal verification for SPARK through two kinds of analyses. The first enables flow analysis to check the initialization of variables, look at data dependencies between inputs and outputs of subprograms, and detect useless code. The second uses deductive verification to generate verification conditions for SMT solvers via a weakest-precondition calculus to detect possible runtime errors and violations of functional contracts.

@@

The verification of the entire functional specification of the CycloneTCP library is beyond the scope of this work. Instead, the aim is to provide significant security hardening to mostly-C libraries by selective use of SPARK. Thus, this work focuses on the hardening of the TCP library areas that its original authors designated as the most vulnerable or crucial to conform to their functional specifications, namely, the Application Programming Interface (API). This mainly falls under two categories: a) proving the conformance to the protocol's functional specification for the translated code, discussed in \Cref{sec:verif}, and b) hardening the user's API to enforce its correct usage by users, discussed in \Cref{sec:API}.

Ada and SPARK offer an easy-to-use mechanism for interfacing with C code. The mechanism allows C-written function to be called by Ada/SPARK code and vice versa and for representing shared data objects. The memory representations of a shared data object must be the same in both the C and the SPARK code. For instance, if a C function is called with a 32-bits integer argument, the SPARK interface of this function should also be called with a 32-bits integer argument. It is the responsibility of the user to ensure this.

@@

In the CycloneTCP library, pointers are used both as the type of variables to pass objects when calling other functions and as the type of structure fields; for instance, in the socket structure to store other complex data structures and recursive pointer-based data structures. SPARK uses a pointer ownership model to prevent bugs involving pointers such as memory leak or double free, which imposes some restrictions on how pointers are being manipulated. When a memory area is allocated, a pointer points to this memory area, and thus, it is considered ``owned.'' The ownership of the memory can be moved, but the memory must be deallocated before the end of the execution of the program.

@@

The socket structure is shared between different functions and contains a large number of fields. Most of the fields are used in library parts that are not considered by this work, and thus, are not relevant for our verification. Thus we use \emph{ghost code} in SPARK to focus the verification effort on the relevant subset of the fields, which also makes the specifications easier to write and understand. This is a solution to the classical problem of specifying the \emph{frame condition} for a function, that is, the set of objects possibly modified by the function. \emph{Ghost code} is used by GNATprove for analysis but is not being compiled in the final executable. More specifically, we define a \emph{model} of the actual socket as a ghost function that can be used in contracts: \spark{Model(Sock)} extracts the fields of interest from socket \spark{Sock}. A common usage is within a function postcondition:

@@

\begin{itemize}
\item The transitions between states must respect the order given by the
state machine described in \Cref{Fig:statemachine}. In particular, a transition from a state to another is valid only if the state machine allows for a transition between these two states.
\item RFC 793 describes the conditions for the state of the socket which need to hold when a user's function is called~\cite[p. 52]{rfc793}, to avoid errors. These conditions were turned into preconditions in our formalization.
\end{itemize}

@@

RFC 793 document provides all the information needed for the allowed transitions between the different TCP possible states. For any transitions directly triggered by calling the TCP user functions, SPARK contracts, representing transitions allowed by the functional specification, are embedded within the SPARK translated code. The helper function \texttt{TCP\_Change\_State} is called explicitly every time a state transition needs to be made to update the socket's state. Any incorrect transition allowed by the code will be detected by GNATprove.

@@

The concurrent implementation of the TCP protocol allows for multiple interactions between the tasks, both synchronous (with a call to \spark{TCP_Wait_For_Events}) or asynchronous (between two user function calls), and thus, numerous possible state changes at any given moment. This makes the functional specification verification significantly challenging, as SPARK does not have a native mode to deal precisely with interactions related to concurrency. To address this, we introduced SPARK functions with contracts that model how the different possible interactions can modify the state of a socket. Finding a correct and precise contract for \spark{TCP_Wait_For_Events} was challenging but needed for the verification of the TCP user functions. The solution is elaborated in the following section.

@@

Symbolic execution is used to verify that all the possible paths when executing the \spark{tcpProcessSegment} function will result in a state that respects the RFC 793 extracted specifications. This process requires three steps shown in \Cref{code:kleedriver}. Firstly, a symbolic incoming segment is created, allowing to account for all possible incoming segments (see lines 1-4). Then at \cref{code:kleedriver:tcpProcessSegment}, function \texttt{tcpProcessSegment} is executed symbolically by KLEE. Finally, the assertion, \texttt{klee\_assert}, at \cref{code:kleedriver:assert} checks if the code leads to a valid state, one that is within TCP specification. If there is no raised assertion, then the contract represented by the assertion is deemed proved, and the relevant C code respects TCP specification regarding that contract. The assertion is then converted to a SPARK contract and added to the function \spark{TCP_Process_One_Segment}. This enables the SPARK prover to reason about the possible side-effects caused by the C-code on the state of a TCP connection and thus, enables the functional correctness verification of the TCP user functions.

@@

One of the most significant problems pointed to by the primary author of the \emph{CycloneTCP} library is the incorrect usage of the library's API. More specifically, the users of the library sometimes call the TCP user functions in the wrong order and forget to check the return code after some calls. This causes their TCP implementation to behave outside of the functional specification of the protocol. The following sections demonstrate how SPARK can be used to enforce a correct usage of the library.

@@

The TCP protocol implies a specific order such that the user can call the TCP user functions without breaking the functional specification of the protocol. This order is also being conveyed by the protocol's state machine given in \Cref{Fig:statemachine}. Ada's preconditions and postconditions are a powerful tool to express such inter-function dependencies, while SPARK technology can be used to guarantee that the assertions always hold. Thus, preconditions and postconditions were introduced to model a partial order on the calls to the TCP user functions.

@@

Simplified contracts extracted from our SPARK implementation are reproduced in \Cref{fig:functionorder}, to demonstrate how SPARK can be used to enforce the correct function calls' ordering. If no error occurs, the function \spark{Socket_Connect} sets \spark{Sock.S_Remote_Ip_Addr} to \spark{Remote_Ip_Addr} which is supposed to be initialized when the function is called. \spark{Socket_Send} requires to be called with a not null socket such that its field \spark{Remote_Ip_Addr} is initialized. The only way to ensure that the precondition hold is to call \spark{Socket_Connect} before \spark{Socket_Send}. By analyzing user's code, GNATprove will statically detect if a call to \spark{Socket_Send} is not preceded by a call to \spark{Socket_Connect}.

@@

These results depend upon other layers in the C programming language, which were not formally verified to the same level. Instead, we used symbolic execution with KLEE to formally verify the contracts used in SPARK to model the corresponding behaviors. The principal function of those layers is to format packets before they are sent, or parse incoming packets, check their integrity, and transmit the resulting payload to the corresponding upper-level layer. This processing part can be a source of errors and bugs as explain in~\cite{Reiher2019RecordFluxFM}. As an alternative to manual translation into SPARK, we envision using the RecordFlux DSL \cite{Reiher2019RecordFluxFM} to specify the parser and printer of the different protocols' packets and generate provable SPARK code from this specification. This would complete our current verification of the CycloneTCP stack towards the lower layers.

>>>>

\author{%
\IEEEauthorblockN{Guillaume Cluzel}
\IEEEauthorblockA{\textit{AdaCore \& ENS de Lyon}}
\and
\IEEEauthorblockN{Kyriakos Georgiou}
\IEEEauthorblockA{\textit{AdaCore \& University of Bristol}}
\and
\IEEEauthorblockN{Yannick Moy}
\IEEEauthorblockA{\textit{AdaCore}}
\and
\IEEEauthorblockN{Clément Zeller}
\IEEEauthorblockA{\textit{Oryx Embedded}}
}

@@

By 2030 it is estimated that one trillion Internet of Things (IoT) devices will be deployed and be part of every aspect of our daily life~\cite{IoT_route:ARM_white_paper}. At the same time, it is evident that security for such systems is a major challenge~\cite{Sec_IoT_Challenges:paper,Cyber_Systems_security:paper}. Any vulnerability in software libraries that are widely used by embedded systems can potentially open the door for compromising the security of billions of deployed IoT devices. Researchers have shown that such vulnerabilities exist: in 2019, critical security vulnerabilities were discovered in a TCP implementation that has been used by billions of IoT devices for more than 13 years~\cite{zero_days_vuln:ARMIS_white_paper}, some of them allowing remote code execution.

Networking communication models define the processes of transferring information from one network component to another. Currently, the two most popular networking communication models are the TCP/IP (Transmission Control Protocol/Internet Protocol) \cite{TCP_IP:ietf_tutorial} and the OSI (Open System Interconnection) \cite{ISO:35.100} models. Layering is an essential design requirement for both models to achieve modularity, flexibility, and abstraction. Applications that only need to use the lower levels' functionality can ignore all the unnecessary upper levels. \Cref{Fig:TcpStack} shows the various layers of the OSI model. Each layer supports a wide range of protocols that are compatible with the layer's specifications. The lowest layers, bottom layers in \Cref{Fig:TcpStack}, are more hardware-oriented, while the top layers are more software-specific. While the \emph{application} level has a wide range of protocols that can be used depending on the application specification, the \emph{transport} layer is heavily dependent on the two dominant protocols: Transmission Control Protocol (TCP) and User Datagram Protocol (UDP). Today, almost every networking application, including IoT devices, is using either the TCP or the UDP protocols to traffic data between the Network and the Session layers. %Thus, the reliability, assurance, and security of these protocols are of paramount importance for critical applications.

@@

In this paper, we take a practical approach, where the security of an existing TCP/IP stack implementation written in the C programming language is incrementally enhanced by replacing parts of its code with formally verified code in SPARK~\cite{mccormick_chapin_2015}, a subset of the Ada programming language supported by formal verification tools. We chose CycloneTCP, a professional-grade open source embedded TCP/IP library developed by the Oryx Embedded company~\cite{CycloneTCP}. \Cref{Fig:TcpStack} shows the protocols supported by the library. This work focuses on the TCP and Socket components of the stack. The library's TCP implementation is meant to conform to the RFC 793 protocol specifications~\cite{rfc793}. The quality of the CycloneTCP is acknowledged by the AMNESIA:33 report~\cite{AMNESIA33}, which classifies it as one of the most resilient TCP/IP stacks.

In total, for this work, 50 functions have been translated from C to SPARK. This accounts for 2266 lines of code, of which 1165 are SPARK logic code (e.g.\ contracts) used to represent and prove TCP specification. The overall proving time with GNATprove from the Community 2021 release is 10 minutes on a 12-processor Intel(R) Core(TM) i7-9750H CPU with 2.60GHz speed and 16GB of RAM. The tools and code are freely available online.

The contributions of this article are the following. First, we show that formal verification can be applied selectively to a layer of an existing TCP/IP library, by translating the TCP layer to SPARK and modelling the lower layers on which it depends. The benefits are three-fold: we found and corrected two bugs in the TCP layer, we proved the correction of that layer, and we defined formal contracts for the upper layers to call the TCP layer. Second, we showed how the models of the lower layers can be validated by symbolic execution with KLEE on the C code of those layers.

@@

TCP is based on a multitasking model. Different tasks can interact and update the socket data structure, used to retain the status of a TCP connection, to handle the various events that can occur within a TCP session. RFC 793, under Section ``3.9. Event Processing'' page 52, describes a possible implementation of how to handle these events based on three tasks: one for the \emph{user calls}, one for the \emph{arriving segments} and one for the \emph{timers}. The role of each of the three tasks can be summarized as follows. \textit{User calls} refer to functions, namely \texttt{open}, \texttt{close}, \texttt{abort}, \texttt{send} and \texttt{receive}, that can be called by the user to control the connection, send, or receive data. These functions can trigger transitions between the connection's states since they are intended to control the connection. In the \textit{arriving segments} task, the received segments are being processed,
and the corresponding messages are sent back. Transitions between states can be triggered on the reception of a segment. Finally, \textit{timers} control the timeouts, for example, the retransmission timeout to retransmit a message or the time-wait timeout to close the connection after a specified amount of time elapsed. Thus, transitions corresponding to the timeout events can also be triggered by this task. A complete description of TCP specification can be found in the RFC 793 document~\cite{rfc793}.

@@

Ada is a general-purpose, procedural and Object-Oriented programming language that puts great emphasis on the safety and correctness of the program. Some of the safety characteristics of the language are its strong typing system and its extensive compile-time and runtime checks. Common C programming vulnerabilities like using unsafe pointers or improper null termination of strings cannot exist in an Ada program, while issues like buffer overflow or integer overflow can be dynamically captured by Ada’s runtime checks and dealt with via exception handlers. Furthermore, Ada provides contracts, such as preconditions and postconditions, as part of the language’s standard syntax. Such contracts are vital for explicitly expressing software and verification requirements in the source code to allow for static analyzers or runtime checks to verify that the stated requirements are met.

By subsetting Ada, the SPARK programming language \cite{mccormick_chapin_2015} provides the largest possible subset of Ada that is suitable for functional specification and static verification. For instance, in SPARK, expressions are free from side effects, and aliasing is forbidden (no two variables can share the exact memory location or overlap in memory), including when using pointers thanks to the use of an ownership policy~\cite{dross2020recursive}.

\emph{GNATprove} \cite{GNATProve:users_manual} is the tool that provides formal verification for SPARK through two kinds of analyses. The first enables flow analysis to check the initialization of variables, verify data dependencies between inputs and outputs of subprograms, and detect useless code. The second uses deductive verification to generate verification conditions for SMT solvers via a weakest-precondition calculus to detect possible runtime errors and violations of functional contracts.

@@

The verification of the entire functional specification of the CycloneTCP library is beyond the scope of this work. Instead, the aim is to provide significant security hardening to mostly-C libraries by selective use of SPARK. Thus, this work focuses on the hardening of the TCP library areas that its original authors designated as the most vulnerable or crucial to conform to their functional specifications, namely, the API. This mainly falls under two categories: a) proving the conformance to the protocol's functional specification for the translated code, discussed in \Cref{sec:verif}, and b) hardening the user's API to enforce its correct usage by users, discussed in \Cref{sec:API}.

Ada and SPARK offer an easy-to-use mechanism for interfacing with C code. The mechanism allows C-written functions to be called by Ada/SPARK code and vice versa and for representing shared data objects. The memory representations of a shared data object must be the same in both the C and the SPARK code. For instance, if a C function is called with a 32-bit integer argument, the SPARK interface of this function should also be called with a 32-bits integer argument. It is the responsibility of the user to ensure this property.

@@

In the CycloneTCP library, pointers are used both as the type of variables passed to functions and as the type of structure fields; for instance, in the socket structure to store other complex data structures and recursive pointer-based data structures. SPARK uses a pointer ownership model to prevent bugs involving pointers such as memory leak or double free, which imposes some restrictions on how pointers are being manipulated. When a memory area is allocated, a pointer points to this memory area, and thus, it is considered ``owned.'' The ownership of the memory can be transferred, but the memory must be deallocated before the end of the execution of the program.

@@

The socket structure is shared between different functions and contains a large number of fields. Most of the fields are used in library parts that are not considered by this work, and thus, are not relevant for our verification. Thus we use \emph{ghost code} in SPARK (that is, special code only meant for verification and not included in the final executable) to focus the verification effort on the relevant subset of the fields, which also makes the specifications easier to write and understand. This is a solution to the classical problem of specifying the \emph{frame condition} for a function, that is, the set of objects possibly modified by the function. \emph{Ghost code} is used by GNATprove for analysis but is not compiled into the final executable. More specifically, we define a \emph{model} of the actual socket as a ghost function that can be used in contracts: \spark{Model(Sock)} extracts the fields of interest from socket \spark{Sock}. A common usage is within a function postcondition:

@@

\begin{itemize}
\item The transitions between states must respect the order given by the
state machine described in \Cref{Fig:statemachine}. In particular, a transition from one state to another is valid only if the state machine allows a transition between these two states.
\item RFC 793 describes the conditions for the state of the socket which need to hold when a user's function is called~\cite[p. 52]{rfc793}, to avoid errors. These conditions were turned into preconditions in our formalization.
\end{itemize}

@@

RFC 793 provides all the information needed for the allowed transitions between the different TCP possible states. For any transitions directly triggered by calling the TCP user functions, SPARK contracts, representing transitions allowed by the functional specification, are embedded within the SPARK translated code. The helper function \texttt{TCP\_Change\_State} is called explicitly every time a state transition needs to be made to update the socket's state. Any incorrect transition allowed by the code will be detected by GNATprove.

@@

The concurrent implementation of the TCP protocol allows for multiple interactions between the tasks, both synchronous (with a call to \spark{TCP_Wait_For_Events}) and asynchronous (between two user function calls), and thus, numerous possible state changes at any given moment. This makes the functional specification verification significantly challenging, as SPARK does not have a native mode to deal precisely with interactions related to concurrency. To address this, we introduced SPARK functions with contracts that model how the different possible interactions can modify the state of a socket. Finding a correct and precise contract for \spark{TCP_Wait_For_Events} was challenging but needed for the verification of the TCP user functions. The solution is elaborated in the following section.

@@

Symbolic execution is used to verify that all the possible paths when executing the \spark{tcpProcessSegment} function will result in a state that respects the RFC 793 extracted specifications. This process requires three steps shown in \Cref{code:kleedriver}. Firstly, a symbolic incoming segment is created, allowing to account for all possible incoming segments (see lines 1-4). Then at \cref{code:kleedriver:tcpProcessSegment}, function \texttt{tcpProcessSegment} is executed symbolically by KLEE. Finally, the assertion, \texttt{klee\_assert}, at \cref{code:kleedriver:assert} checks if the code leads to a valid state, one that is within the TCP specification. If there is no raised assertion, then the contract represented by the assertion is deemed proved, and the relevant C code respects the TCP specification regarding that contract. The assertion is then converted to a SPARK contract and added to the function \spark{TCP_Process_One_Segment}. This enables the SPARK prover to reason about the possible side-effects caused by the C-code on the state of a TCP connection and thus, enables the functional correctness verification of the TCP user functions.

@@

One of the most significant problems pointed to by the primary author of the \emph{CycloneTCP} library is the incorrect usage of the library's API. More specifically, users of the library might call the TCP user functions in the wrong order and forget to check the return code after some calls. This causes their TCP implementation to behave outside of the functional specification of the protocol. The following sections demonstrate how SPARK can be used to enforce a correct usage of the library.

@@

The TCP protocol implies a specific order such that the user can call the TCP user functions without breaking the functional specification of the protocol. This order is also conveyed by the protocol's state machine given in \Cref{Fig:statemachine}. Ada's preconditions and postconditions are a powerful tool to express such inter-function dependencies, while SPARK technology can be used to guarantee that the assertions always hold. Thus, preconditions and postconditions were introduced to model a partial order on the calls to the TCP user functions.

@@

Simplified contracts extracted from our SPARK implementation are reproduced in \Cref{fig:functionorder}, to demonstrate how SPARK can be used to enforce the correct function calls' ordering. If no error occurs, the function \spark{Socket_Connect} sets \spark{Sock.S_Remote_Ip_Addr} to \spark{Remote_Ip_Addr} which is supposed to be initialized when the function is called. \spark{Socket_Send} requires being called with a not null socket such that its field \spark{Remote_Ip_Addr} is initialized. The only way to ensure that the precondition hold is to call \spark{Socket_Connect} before \spark{Socket_Send}. By analyzing the user's code, GNATprove will statically detect if a call to \spark{Socket_Send} is not preceded by a call to \spark{Socket_Connect}.

@@

These results depend upon other layers in the C programming language, which were not formally verified to the same level. Instead, we used symbolic execution with KLEE to formally verify the contracts used in SPARK to model the corresponding behaviors. The principal function of those layers is to format packets before they are sent, or parse incoming packets, check their integrity, and transmit the resulting payload to the corresponding upper-level layer. This processing part can be a source of errors and bugs as explained in~\cite{Reiher2019RecordFluxFM}. As an alternative to manual translation into SPARK, we envision using the RecordFlux DSL \cite{Reiher2019RecordFluxFM} to specify the parser and printer of the different protocols' packets and generate provable SPARK code from this specification. This would complete our current verification of the CycloneTCP stack towards the lower layers.
